\thetitle

### 2020-12-03 sun_phases/spin_qudit_tomo.tex
add acknowledgments
updates and todos given feedback from DB

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By an exact mapping onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary qudit states requires is possible with a minimal number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify numerically for all $d\le200$.
  We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing (a priori) upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating (a posteriori) the statistical error in a reconstructed state.
  Our algorithms motivate a simple randomized tomography protocol, for which we numerically find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
\end{abstract}

@@

Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}.
These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
As a consequence, even well-established techniques can nonetheless be ill-suited for physical platforms with unique or limited capabilities.

@@

The general problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
However, existing protocols either rely infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small faction of the information contained in measurement data.

@@

We begin by introducing a set of qudit operators that are analogous to the spherical harmonics $Y_{\ell m}$, and which play a central role in our work.
Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations \footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
One such basis is that of the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ polarization operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The polarization operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
These properties of the polarization operators allow us to expand any density operator $\rho$ in the polarization operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \rho_{\ell m} T_{\ell m},
  &&
  \rho_{\ell m} \equiv \bk{T_{\ell m}^\dag}_\rho,
  \label{eq:trans_state}
\end{align}
where $\bk{X}_\rho\equiv\tr\p{\rho X}=\obk{\rho|X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The polarization operators can be physically interpreted by the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.

@@

@@

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
  = \delta_{k\ell} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).

@@

Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ by $\bbk{\epsilon_{\v v\ell,0}^2}\le\Gamma_\ell^2$.
The individual bounds on $\bbk{\epsilon_{\v v\ell,0}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
We discuss this maximization problem in Appendix \ref{sec:exact}.
We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
  \label{eq:error}
\end{align}
where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$ is a vector of the degree-$\ell$ polarization operator components $\rho_{\ell m}$ of $\rho$, defined in Eq.~\eqref{eq:trans_state}, and the matrix $\N_{V\ell}$ and vector $\ket{\chi_{V\ell}}$ are defined below.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated a posteriori by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.
After constructing an estimate $\tilde\rho_V$ of $\rho$, the complexity of computing $\E_V\p{\tilde\rho_V}$ is the same as that of $\S_V$ and $\epsilon_V$ (see Figure \ref{fig:times}).

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
After choosing a set of measurement axes $V$, make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates
\begin{align}
  \tilde T_{\v v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
\end{align}
of $\bk{T_{\v v\ell,0}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
Specifically, the $2\ell+1$ (normalized) right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$ and corresponding singular values $M^V_{\ell k}$ of the matrices $M_{V\ell}$ defined in Eq.~\eqref{eq:block} determine the components
\begin{align}
  \tilde\rho^V_{\ell k}
  = \f1{M^V_{\ell k}} \sum_{\v v}
  x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
\end{align}
of
\begin{align}
  \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
  \approx \rho,
  &&
  Q_{\ell k} = \f1{M^V_{\ell k}} \sum_{\v v}
  x^V_{\ell k\v v} T_{\v v\ell,0}.
\end{align}
The total error in $\tilde\rho_V$, measured by its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
We leave the development of tailored and adaptive measurement protocols to future work.

@@

\begin{figure}[hb]
  \centering
  \includegraphics{qudit_axes.pdf}
  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
    Color indicates the qudit dimension $d$ in the legend.
    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.}
  \label{fig:axes}
\end{figure}

@@

@@

Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
We can expand the polarization operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
\begin{align}
  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}
  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell|s\mu},
\end{align}
and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
The covariance between errors in the polarization operator basis is then
\begin{align}
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \delta_{ii'} \times \f1n
  \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \cov_\rho\p{\Pi_{i\mu}, \Pi_{i\mu'}}
  = \delta_{ii'} \times \f1n \cov_\rho\p{T_{i\ell}, T_{i\ell'}},
\end{align}
where we used the fact that the covariance $\cov_\rho\p{X,Y}$ is linear in both $X$ and $Y$.
Due to the appearance of $\delta_{ii'}$ above and the orthogonality of rotated polarization operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, only the variance $\bbk{\epsilon_{i\ell}^2}$ ultimately contributes to reconstruction error.
We therefore seek to find an upper bound on $\bbk{\epsilon_{i\ell}^2}$.

>>>>

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By an exact mapping onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary qudit states requires a minimal number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify numerically for all $d\le200$.
  We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
  Our algorithms motivate a simple randomized tomography protocol, for which we numerically find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
\end{abstract}

@@

Quantum state tomography, the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}.
These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
As a consequence, even well-established techniques can nonetheless be ill-suited for physical platforms with unique or limited capabilities.

@@

The general problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
However, existing protocols either rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small fraction of the information contained in measurement data.

@@

% todo: define spherical harmonics

We begin by introducing a set of qudit operators that are analogous to the spherical harmonic functions $Y_{\ell m}$, and which play a central role in our work.
Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit, distinguishing it from other qudits, is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations \footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
One such basis is that of the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ polarization operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The polarization operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
These properties of the polarization operators allow us to expand any density operator $\rho$ in the polarization operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \rho_{\ell m} T_{\ell m},
  &&
  \rho_{\ell m} \equiv \bk{T_{\ell m}^\dag}_\rho,
  \label{eq:trans_state}
\end{align}
where $\bk{X}_\rho\equiv\tr\p{\rho X}=\obk{\rho|X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The polarization operators can be interpreted in terms of an absorption process, whereby $T_{\ell m}\ket\psi$ is (up to normalization) the state obtained after a spin-$s$ state $\ket\psi$ absorbs a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.

@@

% todo: add reference for Wigner function, clarify / use different language; mention Husimi distribution

% todo: comment further on this representation, motivating signal recovery via the phase-space representation of qudit states

@@

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
  = \delta_{k\ell} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).

@@

Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ by $\bbk{\epsilon_{\v v\ell,0}^2}\le\Gamma_\ell^2$.
The individual bounds on $\bbk{\epsilon_{\v v\ell,0}^2}$ for each axis $\v v$ and degree $\ell$ are tight, but these bounds cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
We discuss this maximization problem in Appendix \ref{sec:exact}.
We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual shot-noise-limited error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
  \label{eq:error}
\end{align}
where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$ is a vector of the degree-$\ell$ polarization operator components $\rho_{\ell m}$ of $\rho$, defined in Eq.~\eqref{eq:trans_state}, and the matrix $\N_{V\ell}$ and vector $\ket{\chi_{V\ell}}$ are defined below.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated a posteriori by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.
After constructing an estimate $\tilde\rho_V$ of $\rho$, the complexity of computing $\E_V\p{\tilde\rho_V}$ is the same as that of $\S_V$ and $\epsilon_V$ (see Figure \ref{fig:times}).

@@

% todo: present protocol as an enumerated list of steps?

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
After choosing a set of measurement axes $V$, make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates
\begin{align}
  \tilde T_{\v v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
\end{align}
of $\bk{T_{\v v\ell,0}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
Specifically, the $2\ell+1$ (normalized) right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$ and corresponding singular values $M^V_{\ell k}$ of the matrices $M_{V\ell}$ defined in Eq.~\eqref{eq:block} determine the components
\begin{align}
  \tilde\rho^V_{\ell k}
  = \f1{M^V_{\ell k}} \sum_{\v v}
  x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
\end{align}
of
\begin{align}
  \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
  \approx \rho,
  &&
  Q_{\ell k} = \f1{M^V_{\ell k}} \sum_{\v v}
  x^V_{\ell k\v v} T_{\v v\ell,0}.
\end{align}
If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

In principle, any information about $\rho$ obtained from prior knowledge or preliminary measurement data can be used to construct tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
We leave the development of tailored and adaptive measurement protocols to future work.

@@

% todo: include table of good measurement axes for specific values of d?
% Yb (5/2), Cr (3), Sr (9/2), Dy (15/2?), Er (19/2)

\begin{figure}[hb]
  \centering
  \includegraphics{qudit_axes.pdf}
  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
    Color indicates the qudit dimension $d$.
    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.}
  \label{fig:axes}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgments}
We thank Anthony M.~Polloreno and Jeremy T.~Young for helpful feedback.
This work is supported  by the DARPA DRINQs grant, the ARO single investigator award W911NF-19-1-0210, AFOSR grant FA9550-19-1-0275, NSF grant PHY-1820885, NSF grant PHY-1734006 (JILA-PFC), and by NIST.
\end{acknowledgments}

@@

Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
We can expand the polarization operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
\begin{align}
  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}
  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell|s\mu},
\end{align}
and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
The covariance between errors in the polarization operator basis is then
\begin{align}
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \delta_{ii'} \times \f1n
  \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \cov_\rho\p{\Pi_{i\mu}, \Pi_{i\mu'}}
  = \delta_{ii'} \times \f1n \cov_\rho\p{T_{i\ell}, T_{i\ell'}},
\end{align}
where we used the fact that the covariance $\cov_\rho\p{X,Y}$ is linear in both $X$ and $Y$.
Due to the appearance of $\delta_{ii'}$ above and the orthogonality of polarization operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, it turns out that only the variances $\bbk{\epsilon_{i\ell}^2}$ will ultimately contribute to reconstruction error (see Appendix \ref{sec:revisiting}).
We therefore seek to find an upper bound on $\bbk{\epsilon_{i\ell}^2}$.
